YouPower -- A Social App for User Engagement in Power Grids

### 2016-10-18 SCITEPRESS_Conference_Latex/designConcept_GP.tex
icnsc 2017 update

Three different levels of energy consumption data are displayed in the app: household, appliances and community. 
At the household level, the current and historical consumptions are displayed. For users with production from renewable sources -- many households from the two Trentino pilot sites have roof-installed PV panels -- the app compares production and consumption levels with the aim to raise awareness of different prosumption patterns and to maximize self-consumption. 
% 
For users with installed smart plugs, the app visualizes consumption patterns at the single appliance level. This is meant to enable users to gain a deeper understanding of the relationship between their daily actions and the resulting energy consumption. 
%
At the community level, aggregated data is displayed on the energy balance of a community. In Italy test sites, a community refers to a whole energy consortium, while in Sweden the housing cooperative a user belongs to. 
The aggregated monthly consumption and production levels are reported. The share of self-consumption is also reported. 
%A single household's consumption level with is compared with the community average. This is meant to represent a powerful tool in helping users make sense of the actual consumption data, driving them towards the goal of a ``fair energy use''. At the same time, this can also be seen as a social mechanism, where community-level dynamics are used as a means to achieve a given goal, in this case increased energy efficiency at the single household level.

\paragraph{Dynamic Time-of-Use Tariffing}
In the Trento pilot site, the main focus is to leverage load elasticity to maximize self-consumption of locally-installed Renewable Energy Sources (RESs). This means that electrical loads shall be shifted to the time of high production. In order to achieve this, the lever identified has been that of price, actuated by means of a dynamic time-of-usage tariffing scheme. 
% 
A predictive engine has been designed to predict the level of production from renewables in the subsequent 72 hours. The engine, which is based on a linear prediction model, uses weather forecast data (solar radiation data) from both public and private sources~ (Meteotrentino \url{http://www.meteotrentino.it}, OpenWeatherMap \url{http://openweathermap.org}, Fondazione Edmund Mach \url{http://www.fmach.it} and US National Weather Service \url{http://nomads.ncep.noaa.gov}) and historical data about the production of single renewable plants provided by local energy consortia.
Estimating consumption patterns based on historical data, a matching engine is designed to forecast whether there will be surplus of local productions. If so, a favorable price is offered to users as an incentive for them to move flexible loads (e.g., dishwater and washing machine) to the perspective time intervals. A higher prices will be used to discourage the use of energy-intensive appliances when local production level is low.
In the app, the energy price is forecasted for the next 72 hours, divided into three-hour intervals. Each interval has a constant price. 
% The indication of the price for each time interval is accompanied by an icon, smiling if the price is below a given threshold (computed based on the historical price of energy in the previous two years) or presenting a sad expression otherwise.

\paragraph{Donation}
A user's contribution towards a better community energy load balance can bring about economic benefits for the electric consortia. 
In both Trento test sites, local generations from RESs exceed consumptions at the aggregated yearly level. Yet there is a timing mismatch so that in order to serve consumption peaks the consortia have to buy (expensive) electricity from the national energy market. At the same time, there are time periods during which production exceeds consumption so that excess energy has to be sold typically at a much lower price on the market.
% 
Thus, the local stakeholders foresee the potentials of donation programmes. 
%The donation mechanism co-designed with the local stakeholder foresees that the economic benefit for the consortia will be partially monetized in order to contribute to a project with a social goal. 
Users in the Trento test sites can optionally sign up for a donation programme organized as a crowdfunding campaign. 
Each donation programme has a well-defined beneficiary. The app gives a description of the programme, information on the current status of the campaign, i.e., how much has been collected with respect to the goal, and plots the trend of donations.

@@

\paragraph{Design Evaluation}
The design of the energy data visualization and comparison part has been validated incrementally with users in the Trentino site. Two focus groups and four workshops were conducted. During the focus group events, ideas about the possible functionalities were used as probes with the participants. During the workshops, practical activities were carried out with the participants in order to develop an idea of the most desirable features and then to receive evaluation feedback.
% 
One issue that emerged was the interest in the production data and the ability to maximize self-consumption. The ideas of the dynamic ToU and the donation programme were also well received.

>>>>

This part of YouPower (see screenshots in Figure~\ref{fig:viz}) is considered for households in the Italian test site.
The main focus is on leveraging load elasticity to maximize self-consumption of locally-installed Renewable Energy Sources (RESs); that is to support the shift of electrical consumption loads to match production peaks. A dynamic time-of-usage signal in combination with a scheme for energy donations is designed to assist this shift. Moreover, in order to make users aware of their consumption patterns, three different levels of energy consumption data are displayed in the app: household, appliances and community.

At the household level, the current and historical consumptions are displayed. For users with production from renewable sources -- many households from the two Trentino test sites have roof-installed PV panels -- the app compares production and consumption levels with the aim to raise awareness of different prosumption patterns. % and to maximize self-consumption. 
% 
For users with installed smart plugs, the app visualizes consumption patterns at the single appliance level. This is meant to enable users to gain a deeper understanding of the relationship between their daily actions and the resulting energy consumption. 
%
At the community level, aggregated data is displayed about the energy balance of a community, (understood as the participants' aggregate consumptions done between consumption under green and red signals) -> ``this is not understandable''.
% The aggregated monthly consumption and production levels are reported. The share of self-consumption is also reported. 
%A single household's consumption level with is compared with the community average. This is meant to represent a powerful tool in helping users make sense of the actual consumption data, driving them towards the goal of a ``fair energy use''. At the same time, this can also be seen as a social mechanism, where community-level dynamics are used as a means to achieve a given goal, in this case increased energy efficiency at the single household level.

\begin{figure*}[t!]
\centering
%\frame{\includegraphics[width=0.7\linewidth]{img/brf.pdf}}
\shadowimage[width=.296\linewidth]{img/viz2.png}
\shadowimage[width=.296\linewidth]{img/viz1.png}
\shadowimage[width=.298\linewidth]{img/viz3.png}
\caption{Energy data part of YouPower}
\label{fig:viz}
\end{figure*}

\paragraph{Dynamic Time-of-Use Signals}

A model (?) is designed to predict the level of production from renewables in the subsequent 72 hours. It is used to generate time-of-use signals that are sent to users. The linear prediction model uses solar radiation data from both public and private sources\footnote{Meteotrentino \url{http://www.meteotrentino.it}, OpenWeatherMap \url{http://openweathermap.org}, Fondazione Edmund Mach \url{http://www.fmach.it} and US National Weather Service \url{http://nomads.ncep.noaa.gov}).} and historical data about the production of renewable plants provided by local energy consortia.
Estimating consumption patterns based on historical data, a matching engine is designed to forecast whether there will be surplus of local productions. If so, a favorable price is offered to users as an incentive for them to move flexible loads (e.g., dishwater and washing machine) to the perspective time intervals. 
In the app, the time-of-use signals are represented by green and red face emoticons in order to identify, respectively, when users can consume and when not. Signals are forecasted
for the following 72 hours, divided into three-hour intervals.

% The indication of the price for each time interval is accompanied by an icon, smiling if the price is below a given threshold (computed based on the historical price of energy in the previous two years) or presenting a sad expression otherwise.

\paragraph{Energy Donation}
A user's contribution towards a better community energy load balance can bring about economic benefits for the electric consortia. Indeed, in both Trento test sites, local generations from RESs can cover the total consumptions at the aggregated yearly level. Yet, there are timing mismatch so that in order to serve demand peaks the consortia have to buy electricity from the national energy grid, while, at other times, local production exceeds demand, so that exceeding energy is sold. From an economic point of view, such transactions with the energy market are unfavorable, because
electricity surpluses are sold at a price that is lower than the national retail market price. At the same time, purchase of electricity from energy market is paid at higher price.
% 
The electrical consortia foresee benefits, in economic and infrastructural terms, from leveraging load shifts and are therefore willing to support an energy donation programme. 
%The donation mechanism co-designed with the local stakeholder foresees that the economic benefit for the consortia will be partially monetized in order to contribute to a project with a social goal. 
Indeed, users in the Trento test sites are part of such a donation programme, which is organized as participatory budgeting process, and from which they can opt-out. KwH consumed during peak of
production (\textit{i.e.} green face emoticon) contribute to the hoarding of collective KwH budget to be allocated to a beneficiary at the end of CIVIS trial period.
In each test site, participants are allowed to submit proposals, in form of a simple and concrete project ideas, in order to be awarded as final beneficiary of the hoarded KwH budget.
The app gives a description of the budgeting programme and provides information about the submitted proposals, so that users can be aware of what kind of beneficiary could benefit from the collective efforts. Furthermore, a simple chart shows how many KwH have been consumed in the green and red time zones.

@@

\paragraph{Design Evaluation}
The design of the Energy Data part has been constructed and validated iteratively with users in Trento test sites. A preliminary inquiry of users' awareness about own energy behaviors and about energy as a collective matter was done through two focus groups. Later, two scenario-based workshops allowed to identify the main users' requirements for supporting the implementation of a load shifting intervention. Finally, two additional workshops were used to validate the proposed ideas, in terms of app's main functionalities, and to produce basic paper-based mock-ups, which CIVIS team used for defining the wire-frame for this part of YouPower.
%
What emerged clearly from the focus groups is that participants share a good sense of energy as a deeply local and collective matter, which is mostly due to the membership-based and cooperative nature of the electrical consortia. They highlighted the relevance of educating people about energy related matters at different levels: from education for the youngest generations in public schools to concrete tools for raising awareness about personal energy behaviors. More concretely, visualization of own consumption (and, for prosumers, production) profiles was considered as a primary and paramount requisite for attempting to shift loads. A predictive system emerged as another cornerstone for enabling participants' flexibility to plan their energy-related behaviors and pursue load shifts. However, comparisons and means for benchmarking consumptions also emerged as highly desirable features for placing people understanding of own energy behaviors into a broader context. The direction towards the participatory budget process for allocating bonus KwH emerged from participants' desire to have a transparent and participated process.
% During the focus group events, ideas about the possible functionalities were used as probes with the participants. During the workshops, practical activities were carried out with the participants in order to develop an idea of the most desirable features and then to receive evaluation feedback.
% 
% One issue that emerged was the interest in the production data and the ability to maximize self-consumption. The ideas of the dynamic ToU and the donation programme were also well received.
